---
title: "Han_spatial_P7_ManualAlignment"
author: "Department of Biostat and Bioinfo"
date: "2024-10-08"
output: 
  html_document:
  theme: journal
  toc: yes
  toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 6 , message = F)
options(future.globals.maxSize = 8000 * 1024^2) # Increase limit to 1000 MiB

library(Seurat) 
library(ggplot2)
library(patchwork)
library(dplyr)
library(dittoSeq)
library(SummarizedExperiment)
library(SpatialExperiment)
library(stringr)

# plan("multisession", workers = 4)   

cutoff.nFeature <- 1000
```


## Data Loading

```{r echo=FALSE}
Sample_A1 <- Load10X_Spatial("sample_A1_C1_manual/outs/", slice = "P7_SampleA")
Sample_B1 <- Load10X_Spatial("sample_B1_C2_manual/outs/", slice = "P7_SampleB")
Sample_A1$orig.ident <- "P7_SampleA"
Sample_B1$orig.ident <- "P7_SampleB"
Sample_B1 <- RenameCells(Sample_B1, new.names = gsub("-1","-2", colnames(Sample_B1)))

Sample_A1
Sample_B1

```

## Merge and filtering

Filter the cells with at least 1000 genes.

```{r}
merge.P7 <- merge(Sample_A1,  y =  Sample_B1, project = "Han", merge.data = FALSE)
merge.P7$orig.ident <- factor(merge.P7$orig.ident, levels = c("P7_SampleA","P7_SampleB"))
Idents(merge.P7) <- "orig.ident"

## filtering  nFeature_Spatial > 1000
merge.P7.filtered <- subset(merge.P7, subset = nFeature_Spatial > cutoff.nFeature) # no quotation mark


VlnPlot(merge.P7.filtered, features = "nCount_Spatial", split.by = "orig.ident")
VlnPlot(merge.P7.filtered, features = "nFeature_Spatial", split.by = "orig.ident")
SpatialFeaturePlot(merge.P7.filtered, features = "nCount_Spatial", ncol = 2, pt.size.factor = 3, alpha = 1) 
SpatialFeaturePlot(merge.P7.filtered, features = "nFeature_Spatial", ncol = 2, pt.size.factor = 3, alpha = 1 ) 


## deleted spots
# merge.P7.deleted <- subset(merge.P7, subset = nFeature_Spatial < cutoff.nFeature) 
# SpatialFeaturePlot(merge.P7.deleted, features = "nCount_Spatial", ncol = 2, pt.size.factor = 1.6) 
# SpatialFeaturePlot(merge.P7.deleted, features = "nFeature_Spatial", ncol = 2, pt.size.factor = 1.6) 

# rm(Sample_A1, Sample_B1, merge.P7)

```


## Prepare for integration

```{r qc, fig.width=12, fig.height=10}
P7.list <- SplitObject(merge.P7.filtered, split.by = "orig.ident")
P7.list <- lapply(X = P7.list, FUN = SCTransform, assay = "Spatial", method = "glmGamPoi", verbose=F)
features <- SelectIntegrationFeatures(object.list = P7.list, nfeatures = 3000)
P7.list <- PrepSCTIntegration(object.list = P7.list, anchor.features = features)

P7.list[["P7_SampleA"]]@images[["P7_SampleB"]] <- NULL
P7.list[["P7_SampleB"]]@images[["P7_SampleA"]] <- NULL

## find anchors
P7.anchor <- FindIntegrationAnchors(object.list = P7.list,
                                         normalization.method = "SCT",
                                         anchor.features = features)

P7.combind.sct <- IntegrateData(anchorset = P7.anchor, normalization.method = "SCT", verbose = F)
rm(P7.anchor, P7.list)
```

## Find neighbors and clusters 

```{r}
resolution <- 1.8
P7.combind.sct  <- RunPCA(P7.combind.sct, verbose = F) %>%
  RunUMAP(reduction = "pca", dims = 1:30) %>%
  FindNeighbors(dims=1:30, verbose = F) %>%
  FindClusters(verbose = F, resolution = resolution)

p.ident <- dittoDimPlot(P7.combind.sct, var = "orig.ident", reduction.use = "umap", 
             do.label = TRUE, labels.highlight = FALSE)
p.clust <- dittoDimPlot(P7.combind.sct, var = "seurat_clusters", reduction.use = "umap", 
             do.label = TRUE, labels.highlight = FALSE)

p.ident | p.clust
```


## Clusters for merged data

```{r echo=FALSE, message=FALSE}
merge.P7.filtered <- SCTransform(merge.P7.filtered, method = "glmGamPoi", assay = "Spatial",
                        verbose = F, return.only.var.genes = FALSE)

merge.P7.filtered  <- RunPCA(merge.P7.filtered, verbose = F) %>%
  RunUMAP(reduction = "pca", dims = 1:30) %>%
  FindNeighbors(dims=1:30, verbose = F) %>%
  FindClusters(verbose = F)

p.ident2 <- dittoDimPlot(merge.P7.filtered, var = "orig.ident", reduction.use = "umap",
             do.label = TRUE, labels.highlight = FALSE)

p.clust2 <- dittoDimPlot(merge.P7.filtered, var = "seurat_clusters", reduction.use = "umap",
             do.label = TRUE, labels.highlight = FALSE)

p.ident2 | p.clust2

dittoDimPlot(merge.P7.filtered, var = "seurat_clusters", reduction.use = "umap",
             do.label = TRUE, labels.highlight = FALSE,
             split.by = "orig.ident", split.ncol = 2)
```

Check whether the order of the barcodes are the same. Replace cell embeddings and seurat clusters from integrated data. 

```{r include=FALSE}
all.equal(colnames(merge.P7.filtered), colnames(P7.combind.sct)) # TRUE 
temp <- match(colnames(merge.P7.filtered), colnames(P7.combind.sct))
all.equal(colnames(merge.P7.filtered), colnames(P7.combind.sct)[temp])

# replace cell embeddings and seurat clusters from integrated data (feature=3000).
merge.P7.filtered@reductions$pca@cell.embeddings <- P7.combind.sct@reductions$pca@cell.embeddings[temp,]
merge.P7.filtered@reductions$umap@cell.embeddings <- P7.combind.sct@reductions$umap@cell.embeddings[temp,]
merge.P7.filtered$seurat_clusters <- P7.combind.sct$seurat_clusters[temp] # check whether is assigned correctly
rm(P7.combind.sct)
```

## UMAP visualization

```{r, fig.width=12, fig.height=6}
p.sample <- dittoDimPlot(merge.P7.filtered, var = "orig.ident", reduction.use = "umap", 
             do.label = FALSE, labels.highlight = FALSE)

p.cluster <- dittoDimPlot(merge.P7.filtered, var = "seurat_clusters", reduction.use = "umap", 
             do.label = TRUE, labels.highlight = TRUE,
             color.panel = dittoColors()[-4])

p.sample.split <- dittoDimPlot(merge.P7.filtered, var = "orig.ident", reduction.use = "umap", 
             do.label = FALSE, labels.highlight = FALSE,
             split.by = "orig.ident", split.ncol = 2)

p.cluster.hightlight <- dittoDimPlot(merge.P7.filtered, var = "seurat_clusters", reduction.use = "umap", 
             do.label = TRUE, labels.highlight = FALSE,
             split.by = "seurat_clusters", split.ncol = 4,
             color.panel = dittoColors()[-4])

p.sample | p.cluster
p.sample.split
# p.cluster.hightlight
```



## Spatial Plot



```{r, fig.width=12, fig.height=8}
pt.size = 2
merge.P7.filtered <- SetIdent(merge.P7.filtered, value = "seurat_clusters")

p1 <- SpatialDimPlot(merge.P7.filtered, pt.size.factor = 1.2) 

p2 <- SpatialDimPlot(merge.P7.filtered, pt.size.factor = pt.size, 
                     cells.highlight = CellsByIdentities(object = merge.P7.filtered, idents = c (13)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 
p3 <- SpatialDimPlot(merge.P7.filtered, pt.size.factor = pt.size,
                     cells.highlight = CellsByIdentities(object = merge.P7.filtered, idents = c (0)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 
p4 <- SpatialDimPlot(merge.P7.filtered, pt.size.factor = pt.size,
                     cells.highlight = CellsByIdentities(object = merge.P7.filtered, idents = c (6)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 

p5 <- SpatialDimPlot(merge.P7.filtered, pt.size.factor = pt.size,
                     cells.highlight = CellsByIdentities(object = merge.P7.filtered, idents = c (15)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 
p6 <- SpatialDimPlot(merge.P7.filtered, pt.size.factor = pt.size,
                     cells.highlight = CellsByIdentities(object = merge.P7.filtered, idents = c (14)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 

p7 <- SpatialDimPlot(merge.P7.filtered, pt.size.factor = pt.size,
                     cells.highlight = CellsByIdentities(object = merge.P7.filtered, idents = c (21)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 


# saveRDS(merge.P7.filtered, "seurat_obj_P7_mergedata.rds")
# ggsave(filename = "sptial_plot_seurat_clusters_P7.png", p1)
# ggsave(filename = "sptial_plot_seurat_clusters_P7.pdf", p1)



p1 
p2
p3 
p4
p5 
p6
p7

## Layers: 13-->0-->6-->15-->14
## 21
```



## Sample distribution in each cluster

```{r, fig.width=10, fig.height=8}
meta.data <- merge.P7.filtered@meta.data
image_coord_A <- merge.P7.filtered@images[["P7_SampleA"]]@coordinates
image_coord_B <- merge.P7.filtered@images[["P7_SampleB"]]@coordinates

image_coord_A$sample_cond <- ifelse(image_coord_A$imagecol < 6550, "InjuryA", "ControlA")
image_coord_B$sample_cond <- ifelse(image_coord_B$imagecol < 6000, "InjuryB", "ControlB")
image_df = rbind(image_coord_A, image_coord_B)

ind <- match(rownames(image_df), rownames(meta.data))
metadata <- cbind(meta.data[ind,], image_df )
metadata$cond <- str_extract(metadata$sample_cond, "Control|Injury")

merge.P7.filtered@meta.data <- metadata
SpatialDimPlot(merge.P7.filtered, group.by = "cond")

p.c.dist <- ggplot(metadata, aes(fill = sample_cond, x=seurat_clusters)) + geom_bar(position = "fill") + scale_fill_brewer()
p1 / p.c.dist
```



## Markers


### All markers
All markers and markers for cluster 5 with `pct.1 > 0.95` and `p_val_adj <= 0.05`

```{r}
# allmarkers <- FindAllMarkers(merge.P7.filtered, assay = "SCT")
# saveRDS(allmarkers, file = "allmarkers_P7_ManualAlign.rds")
# write.csv(allmarkers, file = "allmarkers_P7_ManualAlign.csv")
allmarkers <- readRDS(file = "allmarkers_P7_ManualAlign.rds")
DT::datatable(allmarkers)
```

### Heatmap 

Top 20 over-expressed genes in each cluster with log2FC higher than 0.5

```{r heatmap}
top20 <- allmarkers  %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 0.5) %>%
    top_n(20, avg_log2FC) %>%
    ungroup() 


DoHeatmap(merge.P7.filtered, features = top20$gene) + NoLegend()
```




```{r saverds}
# saveRDS(merge.P7.filtered, file = "seurat_obj_P7_ManualAlign_mergedata_SCT.rds")
merge.P7.filtered <- readRDS("seurat_obj_P7_ManualAlign_mergedata_SCT.rds")
merge.P7.filtered <- SetIdent(merge.P7.filtered, value = "seurat_clusters")
markers.conn1 <- FindMarkers(merge.P7.filtered, ident.1 = 13, ident.2 = 0)
markers.conn2 <- FindMarkers(merge.P7.filtered, ident.1 = 0, ident.2 = 6)
markers.conn3 <- FindMarkers(merge.P7.filtered, ident.1 = 6, ident.2 = 1)
markers.conn4 <- FindMarkers(merge.P7.filtered, ident.1 = 1, ident.2 = 14)

write.csv(markers.conn1, "markers_P7_WithinLayer_conn1.csv")
write.csv(markers.conn2, "markers_P7_WithinLayer_conn2.csv")
write.csv(markers.conn3, "markers_P7_WithinLayer_conn3.csv")
write.csv(markers.conn4, "markers_P7_WithinLayer_conn4.csv")
```


```{r, fig.width=12, fig.height=6}
## GO term
genes <- markers.conn1 %>% filter(avg_log2FC > 0.585, p_val_adj < 0.01) %>% rownames()
GO.res <- enrichGO(genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP",
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")
GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
p1 <- treeplot(GO.res, showCategory = 30)
p1
ggsave(filename = "Treeplot_GOterms_markers_P7_WithinLayer_conn1.png")
write.csv(GO.res@result, file = "GOterms_markers_P7_WithinLayer_conn1.csv")

##
rm(genes, GO.res, p1)
genes <- markers.conn2 %>% filter(avg_log2FC > 0.585, p_val_adj < 0.01) %>% rownames()
GO.res <- enrichGO(genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP",
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")
GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
p1 <- treeplot(GO.res, showCategory = 30)
p1
ggsave(filename = "Treeplot_GOterms_markers_P7_WithinLayer_conn2.png")
write.csv(GO.res@result, file = "GOterms_markers_P7_WithinLayer_conn2.csv")


##
rm(genes, GO.res, p1)
genes <- markers.conn3 %>% filter(avg_log2FC > 0.585, p_val_adj < 0.01) %>% rownames()
GO.res <- enrichGO(genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP",
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")
GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
p1 <- treeplot(GO.res, showCategory = 30)
p1
ggsave(filename = "Treeplot_GOterms_markers_P7_WithinLayer_conn3.png")
write.csv(GO.res@result, file = "GOterms_markers_P7_WithinLayer_conn3.csv")

##
rm(genes, GO.res, p1)
genes <- markers.conn4 %>% filter(avg_log2FC > 0.585, p_val_adj < 0.01) %>% rownames()
GO.res <- enrichGO(genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP",
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")
GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
p1 <- treeplot(GO.res, showCategory = 30)
p1
ggsave(filename = "Treeplot_GOterms_markers_P7_WithinLayer_conn4.png")
write.csv(GO.res@result, file = "GOterms_markers_P7_WithinLayer_conn4.csv")
```




