---
title: "Han_spatial_P1_ManualAlignment"
author: "Department of Biostat and Bioinfo"
date: "2024-10-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 8 , message = F)
options(future.globals.maxSize = 2500 * 1024^2) # Increase limit to 2500 MiB

library(Seurat) 
library(future)
library(ggplot2)
library(patchwork)
library(dplyr)
library(dittoSeq)
library(SummarizedExperiment)
library(SpatialExperiment)
library(stringr)
library(readxl)
library(tidyverse)

# cutoff.nCount <- 1000
cutoff.nFeature <- 1000
```


## Data Loading

```{r}
Sample_A <- Load10X_Spatial("sample_A_manual/outs/", slice = "P1_SampleA")
Sample_B <- Load10X_Spatial("sample_B_manual/outs/", slice = "P1_SampleB")
Sample_A$orig.ident <- "P1_SampleA"
Sample_B$orig.ident <- "P1_SampleB"
Sample_B <- RenameCells(Sample_B, new.names = gsub("-1","-2", colnames(Sample_B)))

Sample_A
Sample_B
```

## Merge and filtering

Filter the cells with at least 1000 gene features. 

```{r}
merge.P1 <- merge(Sample_A,  y =  Sample_B, project = "Han", merge.data = FALSE)
merge.P1$orig.ident <- factor(merge.P1$orig.ident, levels = c("P1_SampleA","P1_SampleB"))
Idents(merge.P1) <- "orig.ident"

## filtering nCount_Spatial > 1000 & nFeature_Spatial > 1000
merge.P1.filtered <- subset(merge.P1, subset =  nFeature_Spatial >= cutoff.nFeature) 
merge.P1.filtered <- SetIdent(merge.P1.filtered, value = "orig.ident")

VlnPlot(merge.P1.filtered, features = "nFeature_Spatial", split.by = "orig.ident")
VlnPlot(merge.P1.filtered, features = "nCount_Spatial", split.by = "orig.ident")

SpatialFeaturePlot(merge.P1.filtered, features = "nFeature_Spatial", ncol = 2, pt.size.factor = 0.8, image.scale = "hires") 
SpatialFeaturePlot(merge.P1.filtered, features = "nCount_Spatial", ncol = 2, pt.size.factor = 0.8, image.scale = "hires") 

## deleted spots
# merge.P1.deleted <- subset(merge.P1, subset = nFeature_Spatial <= cutoff.nFeature)
# SpatialFeaturePlot(merge.P1.deleted, features = "nCount_Spatial", ncol = 2, pt.size.factor = 1.6)
# SpatialFeaturePlot(merge.P1.deleted, features = "nFeature_Spatial", ncol = 2, pt.size.factor = 1.6)

rm(Sample_A, Sample_B, merge.P1)
```


## Prepare for integration

```{r qc, fig.width=12, fig.height=10}
P1.list <- SplitObject(merge.P1.filtered, split.by = "orig.ident")
P1.list <- lapply(X = P1.list, FUN = SCTransform, assay = "Spatial", method = "glmGamPoi", verbose=F)
features <- SelectIntegrationFeatures(object.list = P1.list, nfeatures = 3000)
P1.list <- PrepSCTIntegration(object.list = P1.list, anchor.features = features)

# P1.list[["P1_SampleA"]]@images[["P1_SampleB"]] <- NULL;
# P1.list[["P1_SampleB"]]@images[["P1_SampleA"]] <- NULL

## find anchors
P1.anchor <- FindIntegrationAnchors(object.list = P1.list,
                                    normalization.method = "SCT",
                                    anchor.features = features)

P1.combind.sct <- IntegrateData(anchorset = P1.anchor, normalization.method = "SCT", verbose = F)
rm(P1.anchor, P1.list)
```


## Finding neighbors and clustering

```{r, fig.width=12, fig.height=6}
resolution <- 1.8 # higher resolution, more clusters

P1.combind.sct  <- RunPCA(P1.combind.sct, verbose = F) %>%
  RunUMAP(reduction = "pca", dims = 1:30) %>%
  FindNeighbors(dims=1:30, verbose = F) %>%
  FindClusters(verbose = F, resolution = resolution) # resolution=1.8

p.ident <- dittoDimPlot(P1.combind.sct, var = "orig.ident", reduction.use = "umap", 
             do.label = TRUE, labels.highlight = FALSE)
p.clust <- dittoDimPlot(P1.combind.sct, var = "seurat_clusters", reduction.use = "umap", 
             do.label = TRUE, labels.highlight = FALSE)

p.ident | p.clust
```


## Clusters for merged data

```{r echo=FALSE, message=FALSE, fig.width=12, fig.height=6}
merge.P1.filtered <- SCTransform(merge.P1.filtered, method = "glmGamPoi", assay = "Spatial",
                        verbose = F, return.only.var.genes = FALSE)

merge.P1.filtered  <- RunPCA(merge.P1.filtered, verbose = F) %>%
  RunUMAP(reduction = "pca", dims = 1:30) %>%
  FindNeighbors(dims=1:30, verbose = F) %>%
  FindClusters(verbose = F)

SpatialDimPlot(merge.P1.filtered)
dittoDimPlot(merge.P1.filtered, var = "orig.ident", reduction.use = "umap", 
             do.label = TRUE, labels.highlight = FALSE)

p.ident2 <- dittoDimPlot(merge.P1.filtered, var = "orig.ident", reduction.use = "umap",
             do.label = TRUE, labels.highlight = FALSE)

p.clust2 <- dittoDimPlot(merge.P1.filtered, var = "seurat_clusters", reduction.use = "umap",
             do.label = TRUE, labels.highlight = FALSE)

p.ident2 | p.clust2
# p.ident2 | p.ident

dittoDimPlot(merge.P1.filtered, var = "seurat_clusters", reduction.use = "umap",
             do.label = TRUE, labels.highlight = FALSE,
             split.by = "orig.ident", split.ncol = 2)
# ggsave(filename = "Ditto_Umap_clusters_splitbySample.pdf")
```

Check whether the order of the barcodes are the same. Replace cell embeddings and seurat clusters from integrated data. 

```{r}
all.equal(colnames(merge.P1.filtered), colnames(P1.combind.sct)) # TRUE 
temp <- match(colnames(merge.P1.filtered), colnames(P1.combind.sct))
all.equal(colnames(merge.P1.filtered), colnames(P1.combind.sct)[temp])

# replace cell embeddings and seurat clusters from integrated data (feature=3000).
merge.P1.filtered@reductions$pca@cell.embeddings <- P1.combind.sct@reductions$pca@cell.embeddings[temp,]
merge.P1.filtered@reductions$umap@cell.embeddings <- P1.combind.sct@reductions$umap@cell.embeddings[temp,]

seurat_clusters_merge <- merge.P1.filtered$seurat_clusters
seurat_clusters_integrate <- P1.combind.sct[temp,]$seurat_clusters
seurat_clusters <- data.frame("seurat_clusters_merge" = seurat_clusters_merge, "seurat_clusters_integrate"=seurat_clusters_integrate)
merge.P1.filtered$seurat_clusters <- P1.combind.sct$seurat_clusters[temp] # check whether is assigned correctly
merge.P1.filtered$SCT_label <- merge.P1.filtered$seurat_clusters  ## 25 clusters
rm(P1.combind.sct, seurat_clusters)
```

## UMAP visualization

```{r, fig.width=12, fig.height=6}
p.sample <- dittoDimPlot(merge.P1.filtered, var = "orig.ident", reduction.use = "umap", 
             do.label = FALSE, labels.highlight = FALSE)

p.cluster <- dittoDimPlot(merge.P1.filtered, var = "seurat_clusters", reduction.use = "umap", 
             do.label = TRUE, labels.highlight = TRUE,
             color.panel = dittoColors()[-4])

p.sample.split <- dittoDimPlot(merge.P1.filtered, var = "orig.ident", reduction.use = "umap", 
             do.label = FALSE, labels.highlight = FALSE,
             split.by = "orig.ident", split.ncol = 2)

p.cluster.hightlight <- dittoDimPlot(merge.P1.filtered, var = "seurat_clusters", reduction.use = "umap", 
             do.label = TRUE, labels.highlight = FALSE,
             split.by = "seurat_clusters", split.ncol = 4,
             color.panel = dittoColors()[-4])

p.sample | p.cluster
p.sample.split
```



## Spatial Plot


### Cluster changes from 10--0--1--4--22

```{r echo=FALSE, fig.height=8, fig.width=12}
pt.size = 2.3
merge.P1.filtered <- SetIdent(merge.P1.filtered, value = "seurat_clusters")

p1 <- SpatialDimPlot(merge.P1.filtered, pt.size.factor = pt.size, cols.highlight = c("#DE2D26", "black")) 

p2 <- SpatialDimPlot(merge.P1.filtered, pt.size.factor = pt.size,
                     cells.highlight = CellsByIdentities(object = merge.P1.filtered, idents = c (10)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 
p3 <- SpatialDimPlot(merge.P1.filtered, pt.size.factor = pt.size, 
                     cells.highlight = CellsByIdentities(object = merge.P1.filtered, idents = c (0)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 
p4 <- SpatialDimPlot(merge.P1.filtered, pt.size.factor = pt.size,
                     cells.highlight = CellsByIdentities(object = merge.P1.filtered, idents = c (1)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 

p5 <- SpatialDimPlot(merge.P1.filtered, pt.size.factor = pt.size,
                     cells.highlight = CellsByIdentities(object = merge.P1.filtered, idents = c (4)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 

p6 <- SpatialDimPlot(merge.P1.filtered, pt.size.factor = pt.size,
                     cells.highlight = CellsByIdentities(object = merge.P1.filtered, idents = c (22)),
               facet.highlight = F, cols.highlight = c("#DE2D26", "black")) 

p1
p2 
p3 
p4 
p5 
p6
## resolution = 1.8, cluster layer moves from 10--0--1--4--22

```

## Sample distribution in each cluster

```{r, fig.width=12, fig.height=10}
meta.data <- merge.P1.filtered@meta.data
image_coord_A <- merge.P1.filtered@images[["P1_SampleA"]]@coordinates
image_coord_B <- merge.P1.filtered@images[["P1_SampleB"]]@coordinates

image_coord_A$sample_cond <- ifelse(image_coord_A$imagecol < 9100, "InjuryA", "ControlA")
image_coord_B$sample_cond <- ifelse(image_coord_B$imagecol < -0.33*image_coord_B$imagerow + 10600, 
                                    "InjuryB", "ControlB")
image_df = rbind(image_coord_A, image_coord_B)

ind <- match(rownames(image_df), rownames(meta.data))
metadata <- cbind(meta.data[ind,], image_df )
metadata$cond <- str_extract(metadata$sample_cond, "Control|Injury")

merge.P1.filtered@meta.data <- metadata
SpatialDimPlot(merge.P1.filtered, group.by = "cond")

p.c.dist <- ggplot(metadata, aes(fill = sample_cond, x=seurat_clusters)) + geom_bar(position = "fill") + scale_fill_brewer()
p1 / p.c.dist
```


## Markers


### All markers


```{r allmarkers, echo=FALSE}
# allmarkers <- FindAllMarkers(merge.P1.filtered, assay = "SCT")
# saveRDS(allmarkers, file = "allmarkers_P1_ManualAlign.rds")
# write.csv(allmarkers, file = "allmarkers_P1_ManualAlign.csv")

allmarkers <- readRDS(file = "allmarkers_P1_ManualAlign.rds")
DT::datatable(allmarkers)
```


### Heatmap 

Top 20 over-expressed genes in each cluster with log2FC higher than 0.5

```{r heatmap, echo=FALSE}
top20 <- allmarkers  %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 0.5) %>%
    top_n(20, avg_log2FC) %>%
    ungroup() 

DoHeatmap(merge.P1.filtered, features = top20$gene) + NoLegend()
```



```{r}
saveRDS(merge.P1.filtered, "seurat_obj_P1_ManualAlign_mergedata_SCT.rds")
```


## Markers between layers

```{r}
merge.P1.filtered <- readRDS("seurat_obj_P1_ManualAlign_mergedata_SCT.rds")
markers.conn1 <- FindMarkers(merge.P1.filtered, ident.1 = 10, ident.2 = 0)
markers.conn2 <- FindMarkers(merge.P1.filtered, ident.1 = 0, ident.2 = 1)
markers.conn3 <- FindMarkers(merge.P1.filtered, ident.1 = 1, ident.2 = 4)
markers.conn4 <- FindMarkers(merge.P1.filtered, ident.1 = 4, ident.2 = 22)

write.csv(markers.conn1, "markers_P1_WithinLayer_conn1.csv")
write.csv(markers.conn2, "markers_P1_WithinLayer_conn2.csv")
write.csv(markers.conn3, "markers_P1_WithinLayer_conn3.csv")
write.csv(markers.conn4, "markers_P1_WithinLayer_conn4.csv")
```




```{r, fig.width=12, fig.height=6}
## GO term
genes <- markers.conn1 %>% filter(avg_log2FC > 0.585, p_val_adj < 0.01) %>% rownames()
GO.res <- enrichGO(genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP",
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")
GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
p1 <- treeplot(GO.res, showCategory = 30)
p1
ggsave(filename = "Treeplot_GOterms_markers_P1_WithinLayer_conn1.png")
write.csv(GO.res@result, file = "GOterms_markers_P1_WithinLayer_conn1.csv")

##
rm(genes, GO.res, p1)
genes <- markers.conn2 %>% filter(avg_log2FC > 0.585, p_val_adj < 0.01) %>% rownames()
GO.res <- enrichGO(genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP",
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")
GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
p1 <- treeplot(GO.res, showCategory = 30)
p1
ggsave(filename = "Treeplot_GOterms_markers_P1_WithinLayer_conn2.png")
write.csv(GO.res@result, file = "GOterms_markers_P1_WithinLayer_conn2.csv")

##
rm(genes, GO.res, p1)
genes <- markers.conn3 %>% filter(avg_log2FC > 0.585, p_val_adj < 0.01) %>% rownames()
GO.res <- enrichGO(genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP",
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")
GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
p1 <- treeplot(GO.res, showCategory = 30)
p1
ggsave(filename = "Treeplot_GOterms_markers_P1_WithinLayer_conn3.png")
write.csv(GO.res@result, file = "GOterms_markers_P1_WithinLayer_conn3.csv")

##
rm(genes, GO.res, p1)
genes <- markers.conn4 %>% filter(avg_log2FC > 0.585, p_val_adj < 0.01) %>% rownames()
GO.res <- enrichGO(genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP",
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")
GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
p1 <- treeplot(GO.res, showCategory = 30)
p1
ggsave(filename = "Treeplot_GOterms_markers_P1_WithinLayer_conn4.png")
write.csv(GO.res@result, file = "GOterms_markers_P1_WithinLayer_conn4.csv")
```

```{r}
merge.P1.filtered <- readRDS("seurat_obj_P1_ManualAlign_mergedata_SCT.rds")
table(merge.P1.filtered$tissue)
SHAM <- subset(merge.P1.filtered, subset = cond == "Control")
SHAM <- SetIdent(SHAM, value = "tissue")
Feature_avgExp_sham <- AverageExpression(SHAM, assays = "SCT", return.seurat = F)$SCT
Feature_avgExp_sham <- as.data.frame(Feature_avgExp_sham)
Feature_avgExp_sham_order <- Feature_avgExp_sham %>% arrange(desc(all)) %>% head(n=3000) %>% rownames() 


LAD <- subset(merge.P1.filtered, subset = cond == "Injury")
LAD <- SetIdent(LAD, value = "tissue")
Feature_avgExp_LAD <- AverageExpression(LAD, assays = "SCT", return.seurat = F)$SCT
Feature_avgExp_LAD <- as.data.frame(Feature_avgExp_LAD)
Feature_avgExp_LAD_order <- Feature_avgExp_LAD %>% arrange(desc(all)) %>% head(n=3000) %>% rownames() 

highly_express_genes <- c(Feature_avgExp_sham_order, Feature_avgExp_LAD_order) %>% unique()
write.csv(highly_express_genes, file = "Merged_Top3000_highlyexpGenes_P1.csv")

```


