---
title: "scRNA integration across timepoints, samples and condtions"
author: "Department of Biostat and Bioinfo"
date: "2025-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 6 , message = F, warning = F)
options(future.globals.maxSize = 5000 * 1024^2) # Increase limit to 
# options(scipen = 999)
library(Seurat) 
library(ggplot2)
library(patchwork)
library(dplyr)
library(SummarizedExperiment)
library(tidyverse)
library(ggfortify)
library(scales)
library(data.table)
library(SPOTlight)
library(viridis)
library(DT)
library(RColorBrewer)
library(MAESTRO)
library(clusterProfiler)
library(enrichplot)

c25 <- c(
  "dodgerblue2", "#E31A1C", # red
  "green4",
  "#6A3D9A", # purple
  "#FF7F00", # orange
  "black", "gold1",
  "skyblue2", "#FB9A99", # lt pink
  "palegreen2",
  "#CAB2D6", # lt purple
  "#FDBF6F", # lt orange
  "gray70", "khaki2",
  "maroon", "orchid1", "deeppink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)

topn = 3000
bkg_genes <- read.csv(file = paste0("../GOterms/Top",topn,"_highlyexpGenes_all.csv") )
bkg_genes <- bkg_genes$x

msig.mhallmark <- read.gmt("~/Downloads/myDB/annotation/MSigDB_mouse_2024/mh.all.v2024.1.Mm.symbols.gmt")
msig.m2 <- read.gmt("~/Downloads/myDB/annotation/MSigDB_mouse_2024/m2.cp.v2024.1.Mm.symbols.gmt")
```

 
## Data input and split 

Input rds object from timepoint P1 and P7, split the object based on sample_condition to get 8 replicates.


```{r fig.height=6, fig.width=12}
## subset time_sample_cond, get 8 rep

merge.P1.filtered <- readRDS("~/Downloads/Projects/Han_spatial_2024/P1/seurat_obj_P1_ManualAlign_mergedata_SCT.rds")
# merge.P1.filtered@assays$SCT <-  NULL
P1_sampleA_Inj <- subset(merge.P1.filtered, subset = sample_cond=="InjuryA" )
P1_sampleA_Ctr <- subset(merge.P1.filtered, subset = sample_cond=="ControlA" )
P1_sampleB_Inj <- subset(merge.P1.filtered, subset = sample_cond=="InjuryB" )
P1_sampleB_Ctr <- subset(merge.P1.filtered, subset = sample_cond=="ControlB" ); rm(merge.P1.filtered)

merge.P7.filtered <- readRDS("~/Downloads/Projects/Han_spatial_2024/P7/seurat_obj_P7_ManualAlign_mergedata_SCT.rds")
# merge.P7.filtered@assays$SCT <-  NULL
P7_sampleA_Inj <- subset(merge.P7.filtered, subset = sample_cond=="InjuryA" )
P7_sampleA_Ctr <- subset(merge.P7.filtered, subset = sample_cond=="ControlA" )
P7_sampleB_Inj <- subset(merge.P7.filtered, subset = sample_cond=="InjuryB" )
P7_sampleB_Ctr <- subset(merge.P7.filtered, subset = sample_cond=="ControlB" ); rm(merge.P7.filtered)


## Simple merging

merge.postnatal <- merge(P1_sampleA_Inj, y = c(P1_sampleA_Ctr, P1_sampleB_Inj, P1_sampleB_Ctr,
                                         P7_sampleA_Inj, P7_sampleA_Ctr, P7_sampleB_Inj, P7_sampleB_Ctr) 
                   )

rm(P1_sampleA_Inj, P1_sampleA_Ctr, P1_sampleB_Inj, P1_sampleB_Ctr,
    P7_sampleA_Inj, P7_sampleA_Ctr, P7_sampleB_Inj, P7_sampleB_Ctr)

merge.postnatal$timesamplecond <- paste0(merge.postnatal$orig.ident,"_", merge.postnatal$cond)
merge.postnatal$time <- substr(merge.postnatal$orig.ident, 1, 2)
```


## Integration

Split the Spatial assay according to its meta timesamolecond, and normalize it using SCTransform.
In integration step, use CCA method in function IntegrateLayers. Use resolution 1.2 to cluster the spots.


```{r integrate, fig.height=6, fig.width=12, warning=FALSE}

## prestep
merge.postnatal[["Spatial"]] <- split(merge.postnatal[["Spatial"]], f = merge.postnatal$timesamplecond)
merge.postnatal <- SCTransform(merge.postnatal, assay = "Spatial")
merge.postnatal <- RunPCA(merge.postnatal)
merge.postnatal <- FindNeighbors(merge.postnatal, dims = 1:30, reduction = "pca")
merge.postnatal <- FindClusters(merge.postnatal, cluster.name = "unintegrated_clusters")
# merge.postnatal <- RunUMAP(merge.postnatal, dims = 1:30)

p1 <- DimPlot(merge.postnatal, group.by = "time") 
p2 <- DimPlot(merge.postnatal, group.by = "sample_cond") 
p3 <- DimPlot(merge.postnatal, group.by = "timesamplecond")

p1 | p2 | p3



####==== CCA integration ====####
merge.postnatal <- IntegrateLayers(object = merge.postnatal, method = CCAIntegration,
                              verbose = F, normalization.method = "SCT")
merge.postnatal[["Spatial"]] <- JoinLayers(merge.postnatal[["Spatial"]])

####==== Harmony ====####
# merge.postnatal <- IntegrateLayers(object = merge.postnatal, method = HarmonyIntegration, 
#                               verbose = F, normalization.method = "SCT") 


merge.postnatal <- FindNeighbors(merge.postnatal, reduction = "integrated.dr", dims = 1:30)
merge.postnatal <- FindClusters(merge.postnatal, resolution = 1.2)
merge.postnatal <- RunUMAP(merge.postnatal, dims = 1:30, reduction = "integrated.dr")


DimPlot(merge.postnatal, reduction = "umap", group.by = c("seurat_clusters", "orig.ident"), cols = c25, pt.size = 0.6)
DimPlot(merge.postnatal, reduction = "umap", group.by = c("orig.ident", "sample_cond", "cond"), cols = c25, pt.size = 0.6)
DimPlot(merge.postnatal, reduction = "umap", group.by = c("sample_cond",  "timesamplecond"), 
        cols = brewer.pal(10,"Paired"), pt.size = 0.6)

```


```{r fig.width=12, fig.height=8}
table(merge.postnatal$seurat_clusters, merge.postnatal$timesamplecond)
table(merge.postnatal$seurat_clusters, merge.postnatal$sample_cond)

ptsize <- 4
immune.marker <- c("Apoe", "Nppa", "C1qa", "Spp1")

FeaturePlot(merge.postnatal, features = immune.marker, split.by = "orig.ident")
SpatialFeaturePlot(merge.postnatal, features = immune.marker, pt.size.factor = ptsize )

SpatialDimPlot(merge.postnatal, pt.size.factor = ptsize, ncol = 2) 
SpatialDimPlot(merge.postnatal, pt.size.factor = ptsize, ncol = 4,   cols = c("4" = "#DE2D26", "6" = "green", 
                                                                              "8" = "yellow",  "9" = "blue")) 
```

## DEG

```{r}
for (i in 1:length(merge.postnatal[["SCT"]]@SCTModel.list)){
  slot(merge.postnatal@assays$SCT@SCTModel.list[[i]], name = "umi.assay") <- "Spatial"
}

merge.postnatal <- PrepSCTFindMarkers(merge.postnatal, assay = "SCT", verbose = F)
all_markers <- FindAllMarkers(merge.postnatal, verbose = F)
```


## Find DEGs between Injury and Control for some clusters 


```{r}
logfc_cutoff <- 1
padj_cutoff <- 0.01

merge.postnatal.P1 <- subset(merge.postnatal, subset=time =="P1")
merge.postnatal.P7 <- subset(merge.postnatal, subset=time =="P7")

```



### Cluster 4: P1

```{r warning=FALSE}

DEG_comp_cond <- FindMarkers(merge.postnatal.P1, group.by = "cond", ident.1 = "Injury", ident.2 = "Control", 
                             subset.ident = "4", recorrect_umi = FALSE)
datatable(DEG_comp_cond)


genes <- DEG_comp_cond %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
overlap_genes

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p1

}

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result, extensions="Buttons",
            options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### Cluster 4: P7

```{r warning=FALSE}

DEG_comp_cond <- FindMarkers(merge.postnatal.P7, group.by = "cond", ident.1 = "Injury", ident.2 = "Control", 
                             subset.ident = "4", recorrect_umi = FALSE)
datatable(DEG_comp_cond)


genes <- DEG_comp_cond %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
overlap_genes

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p1

}

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result, extensions="Buttons",
            options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### Cluster 6: P1

```{r warning=FALSE}
rm(DEG_comp_cond, genes, GO.res, p1, p2, ER.h, ER.cp)
DEG_comp_cond <- FindMarkers(merge.postnatal.P1, group.by = "cond", ident.1 = "Injury", ident.2 = "Control", 
                             subset.ident = "6", recorrect_umi = FALSE)
datatable(DEG_comp_cond)


genes <- DEG_comp_cond %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
overlap_genes
GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p1

}

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result, extensions="Buttons",
            options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### Cluster 6: P7

```{r warning=FALSE}
rm(DEG_comp_cond, genes, GO.res, p1, p2, ER.h, ER.cp)
DEG_comp_cond <- FindMarkers(merge.postnatal.P7, group.by = "cond", ident.1 = "Injury", ident.2 = "Control", 
                             subset.ident = "6", recorrect_umi = FALSE)
datatable(DEG_comp_cond)


genes <- DEG_comp_cond %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
overlap_genes
GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p1

}

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result, extensions="Buttons",
            options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### Cluster 8: P1

```{r warning=FALSE}
rm(DEG_comp_cond, genes, GO.res, p1, p2, ER.h, ER.cp)
DEG_comp_cond <- FindMarkers(merge.postnatal.P1, group.by = "cond", ident.1 = "Injury", ident.2 = "Control", 
                             subset.ident = "8", recorrect_umi = FALSE)
datatable(DEG_comp_cond)


genes <- DEG_comp_cond %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
overlap_genes
GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p1

}

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result, extensions="Buttons",
            options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```



### Cluster 8: P7

```{r warning=FALSE}
rm(DEG_comp_cond, genes, GO.res, p1, p2, ER.h, ER.cp)
DEG_comp_cond <- FindMarkers(merge.postnatal.P7, group.by = "cond", ident.1 = "Injury", ident.2 = "Control", 
                             subset.ident = "8", recorrect_umi = FALSE)
datatable(DEG_comp_cond)


genes <- DEG_comp_cond %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
overlap_genes
GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p1

}

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result, extensions="Buttons",
            options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### Cluster 9: P1

```{r warning=FALSE}
rm(DEG_comp_cond, genes, GO.res, p1, p2, ER.h, ER.cp)
DEG_comp_cond <- FindMarkers(merge.postnatal.P1, group.by = "cond", ident.1 = "Injury", ident.2 = "Control", 
                             subset.ident = "9", recorrect_umi = FALSE)
datatable(DEG_comp_cond)


genes <- DEG_comp_cond %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
overlap_genes

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p1

}

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result, extensions="Buttons",
            options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```







### Cluster 9: P7

```{r warning=FALSE}
rm(DEG_comp_cond, genes, GO.res, p1, p2, ER.h, ER.cp)
DEG_comp_cond <- FindMarkers(merge.postnatal.P7, group.by = "cond", ident.1 = "Injury", ident.2 = "Control", 
                             subset.ident = "9", recorrect_umi = FALSE)
datatable(DEG_comp_cond)


genes <- DEG_comp_cond %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
overlap_genes

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p1

}

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result, extensions="Buttons",
            options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```



## Find DEGs between P1 and P7 for sample condition

```{r}
merge.postnatal.Inj <- subset(merge.postnatal, subset=cond =="Injury")
merge.postnatal.Ctr <- subset(merge.postnatal, subset=cond =="Control")
```


### Cluster 4: Injury

```{r warning=FALSE}
rm(genes, GO.res, p1, p2, ER.h, ER.cp)

DEG_comptime <- FindMarkers(merge.postnatal.Inj, group.by = "time", ident.1 = "P1", ident.2 = "P7", 
                             subset.ident = "4", recorrect_umi = FALSE)

genes <- DEG_comptime %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
# overlap_genes
GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)
GO.res <- pairwise_termsim(GO.res, method = "JC")

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2

datatable(GO.res@result)
```



### Cluster 4: Control

```{r warning=FALSE}
rm(DEG_comptime, genes, GO.res, p1, p2, ER.h, ER.cp)

DEG_comptime <- FindMarkers(merge.postnatal.Ctr, group.by = "time", ident.1 = "P1", ident.2 = "P7", 
                             subset.ident = "4", recorrect_umi = FALSE)

genes <- DEG_comptime %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
# overlap_genes
GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)
n.GOterm <- sum(GO.res@result$p.adjust <=0.05)
if (n.GOterm > 0) GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")

  p1
  p2
  datatable(GO.res@result)
}


```


### Cluster 6: Injury

```{r warning=FALSE}
rm(DEG_comptime, genes, GO.res, p1, p2, ER.h, ER.cp)

DEG_comptime <- FindMarkers(merge.postnatal.Inj, group.by = "time", ident.1 = "P1", ident.2 = "P7", 
                             subset.ident = "6", recorrect_umi = FALSE)

genes <- DEG_comptime %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
# overlap_genes
GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)
n.GOterm <- sum(GO.res@result$p.adjust <=0.05)
if (n.GOterm > 0) GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")

  p1
  p2
  datatable(GO.res@result)
}

```



### Cluster 8: Injury

```{r warning=FALSE}
rm(DEG_comptime, genes, GO.res, p1, p2, ER.h, ER.cp)

DEG_comptime <- FindMarkers(merge.postnatal.Inj, group.by = "time", ident.1 = "P1", ident.2 = "P7", 
                             subset.ident = "8", recorrect_umi = FALSE)

genes <- DEG_comptime %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
# overlap_genes
GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)
n.GOterm <- sum(GO.res@result$p.adjust <=0.05)
if (n.GOterm > 0) GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
 
  p1 <- treeplot(GO.res, showCategory = 30)
  p1
  
}

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result)
```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```



### Cluster 8: Control

```{r warning=FALSE}
rm(DEG_comptime, genes, GO.res, p1, p2, ER.h, ER.cp)

DEG_comptime <- FindMarkers(merge.postnatal.Ctr, group.by = "time", ident.1 = "P1", ident.2 = "P7", 
                             subset.ident = "8", recorrect_umi = FALSE)

genes <- DEG_comptime %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)
overlap_genes
GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)
n.GOterm <- sum(GO.res@result$p.adjust <=0.05)
if (n.GOterm > 0) GO.res <- pairwise_termsim(GO.res, method = "JC")


p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result)


```


### Cluster 9: Injury

```{r warning=FALSE}
rm(DEG_comptime, genes, GO.res, p1, p2, ER.h, ER.cp)


DEG_comptime <- FindMarkers(merge.postnatal.Inj, group.by = "time", ident.1 = "P1", ident.2 = "P7", 
                             subset.ident = "9", recorrect_umi = FALSE)

genes <- DEG_comptime %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.GOterm <- sum(GO.res@result$p.adjust <=0.05)
if (n.GOterm > 0) GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
 
  p1 <- treeplot(GO.res, showCategory = 30)
  p1
  
}
p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2

datatable(GO.res@result)
```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### Cluster 9: Control

```{r warning=FALSE}
rm(DEG_comptime, genes, GO.res, p1, p2, ER.h, ER.cp)

DEG_comptime <- FindMarkers(merge.postnatal.Ctr, group.by = "time", ident.1 = "P1", ident.2 = "P7", 
                             subset.ident = "4", recorrect_umi = FALSE)

genes <- DEG_comptime %>% filter(avg_log2FC > logfc_cutoff, p_val_adj < padj_cutoff) %>% rownames()
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)
n.GOterm <- sum(GO.res@result$p.adjust <=0.05)
if (n.GOterm > 0) GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
 
  p1 <- treeplot(GO.res, showCategory = 30)
  p1
  
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  p2

  datatable(GO.res@result)

  
}

```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

## Cell type annotation

```{r, fig.width=12, fig.height=6}
data("mouse.all.droplet.TabulaMuris")
sig.mouse.heart <- mouse.all.droplet.TabulaMuris[grep("\\(Heart_and_Aorta\\)", mouse.all.droplet.TabulaMuris$cluster) ,]
sig.mouse.heart$cluster <- gsub("\\(Heart_and_Aorta\\)", "", sig.mouse.heart$cluster)
rownames(all_markers) <- NULL
colnames(all_markers) <- c("p_val","avg_logFC","pct.1","pct.2","p_val_adj","cluster","gene")

merge.postnatal <- RNAAnnotateCelltype(RNA= merge.postnatal, genes = all_markers, 
                                    signatures = sig.mouse.heart)

table(merge.postnatal$timesamplecond, merge.postnatal$assign.ident)
DimPlot(merge.postnatal, group.by = "assign.ident", label = T)
dittoSeq::dittoDimPlot(merge.postnatal, var ="assign.ident" , split.by   = "timesamplecond")
```





```{r eval=FALSE, include=FALSE}
cls = 0
cls_topmarkers <- all_markers %>% filter(avg_logFC > 0 , cluster == cls) %>% pull(gene) %>% head(n=25)
overlaps_tbl <- clustermole::clustermole_overlaps(genes = cls_topmarkers, species = "mm")
overlaps_tbl %>% filter(species == "Mouse", fdr <=0.05)


cls = 1
cls_topmarkers <- all_markers %>% filter(avg_logFC > 0 , cluster == cls) %>% pull(gene) %>% head(n=25)
overlaps_tbl <- clustermole::clustermole_overlaps(genes = cls_topmarkers, species = "mm")
overlaps_tbl %>% filter(species == "Mouse", fdr <=0.05)

cls = 6
cls_topmarkers <- all_markers %>% filter(avg_logFC > 0 , cluster == cls) %>% pull(gene) %>% head(n=25)
overlaps_tbl <- clustermole::clustermole_overlaps(genes = cls_topmarkers, species = "mm")
overlaps_tbl %>% filter(species == "Mouse", fdr <=0.05)

cls = 8
cls_topmarkers <- all_markers %>% filter(avg_logFC > 0 , cluster == cls) %>% pull(gene) %>% head(n=25)
overlaps_tbl <- clustermole::clustermole_overlaps(genes = cls_topmarkers, species = "mm")
overlaps_tbl %>% filter(species == "Mouse")
```


## GO terms and MsigDB pathway

### DEGs in cluster 0

```{r, fig.width=15, fig.height=8}
all_pos_markers <- all_markers %>% filter(avg_logFC > 0)

cls <- 0
genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", 
                   universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

GO.res <- pairwise_termsim(GO.res, method = "JC")

if (dim(GO.res@termsim)[1] >=5 ){
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  
  p1 <- treeplot(GO.res, showCategory = 30)
  p1

}

p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
p2
datatable(GO.res@result, extensions="Buttons",
            options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### DEGs in cluster 1

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 1

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))
  }

```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### DEGs in cluster 2

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 2

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### DEGs in cluster 3

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 3

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }

```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### DEGs in cluster 4

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 4

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### DEGs in cluster 5

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 5

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### DEGs in cluster 6

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 6

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### DEGs in cluster 7

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 7

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```

```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### DEGs in cluster 8

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 8

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.05, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### DEGs in cluster 9

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 9

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```

### DEGs in cluster 10

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 10

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### DEGs in cluster 11

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 11

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### DEGs in cluster 12

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 12

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### DEGs in cluster 13

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 13

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### DEGs in cluster 14

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 14

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### DEGs in cluster 15

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 15

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```


### DEGs in cluster 16

```{r, fig.width=12, fig.height=8}
rm(genes, overlap_genes, GO.res,  p1, p2)
cls <- 16

genes <- all_pos_markers %>% filter(cluster == cls) %>% pull(gene) 
overlap_genes <- intersect(genes, bkg_genes)

GO.res <- enrichGO(overlap_genes, OrgDb = "org.Mm.eg.db", keyType = "SYMBOL", ont = "BP", universe = bkg_genes,
                   pvalueCutoff = 0.05, qvalueCutoff = 0.2, readable=T)

n.enrich <- GO.res@result %>% filter(p.adjust <=0.05, qvalue <= 0.2) %>% nrow()

if (n.enrich >=5){
  GO.res <- pairwise_termsim(GO.res, method = "JC")
  GO.res <- simplify(GO.res, cutoff=0.7, by="p.adjust", select_fun=min)
  p1 <- treeplot(GO.res, cluster.params=list(n=min(n.enrich, 5)) )
  print(p1)
  p2 <- mutate(GO.res, qscore = -log(p.adjust, base=10)) %>% barplot(x="qscore")
  print(p2)
  datatable(GO.res@result, extensions="Buttons",
          options = list(dom="Blfrtip", buttons=c("copy","csv","excel","pdf","print")))

  }
```


```{r, fig.width=12, fig.height=6}
ER.h <- enricher(overlap_genes, TERM2GENE = msig.mhallmark, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Hallmarks"

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  dotplot(ER.h, title=pathway.name)
}

if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  barplot(ER.h, title = pathway.name)
}
if ( !is.null(ER.h) && nrow(ER.h) > 0 ) {
  datatable(ER.h@result)
}

```


```{r, fig.width=12, fig.height=6}
ER.cp <- enricher(overlap_genes, TERM2GENE = msig.m2, pvalueCutoff  = 0.01, qvalueCutoff = 0.05, 
               pAdjustMethod = "BH", minGSSize = 10, maxGSSize = 500, universe = bkg_genes)

pathway.name <- "Canonical pathways"

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  dotplot(ER.cp, title=pathway.name)
}

if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  barplot(ER.cp, title = pathway.name)
}
if ( !is.null(ER.cp) && nrow(ER.cp) > 0 ) {
  datatable(ER.cp@result)
}

```




```{r include=FALSE}
saveRDS(merge.postnatal, file = "Postnatal_P1P7_integration.rds")
saveRDS(all_markers, "markers_for_merged_postnatal.rds")


# merge.postnatal <- readRDS("Postnatal_P1P7_integration.rds")
# all_markers <- readRDS("markers_for_merged_postnatal.rds")
```






